{CPU-Scheduling}

### 2024-12-31 LaTeX/Sources/Advanced.tex
added cfs and conclusion

>>>>

\section{THe Linux 2.6 Completely Fair Scheduler}

As of 2023, this scheduler was replaced in the Linux kernel 6.6 by another scheduler based on the \emph{EEVDF}.
The CFS is therefore technically retired, however it was used for a very long time, more than 15 years in fact.
The goal of the policy was to be fair. To do this it was based on a proportional-share algorithm.

\subsection{How it works}

The aim is to divide the CPU time completely fairly.
This means that each process gets a $\frac{100\%}{N}$ share of runtime, where $N$ is the total number of jobs competing.
The policy turns out to be periodic and in each cycle everyone runs the same amount of time. 
This is the ideal outcome of course.
To do this we measure a so called vruntime.
This is pretty similar to the pass value in stride scheduling, with the main difference being how this value is incremented.
While in stride scheduling we use strides, in the CFS we increment the vruntime by the time that the process ran.
If we have to make a scheduling decision, we just choose the process with the smallest vruntime.
New jobs just receive the lowest vruntime value, so that they get a "priority" in the beginning.

\subsection{Red-Black Binary Trees}

Here I will quickly mention, why a red black binary tree is used, but not how it is implemented.
That concept is out of the scope of this paper, because the goal is to explain the scheduling policies and not data structures.
The CFS uses a special Red-Black Binary tree as its data structure.
The Red-Black Binary tree is a Balanced Binary Search tree.
This means that it is sorted from lowest vruntime to highest vruntime.
In addition to that the smallest branch is no smaller than half the size of the longest branch.
The lowest vruntime is on the left of the tree.
This is useful, because it is really easy to find the next job to run: It is on the outer most left leaf.

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{Assets/Red-Black-Tree.png}
    \caption{Red-Black Binary Tree\\Source: \url{https://en.wikipedia.org/wiki/Red\%E2\%80\%93black_tree}}
    \label{fig:stride-scheduling}
\end{figure}

After the process finished running, we just need to put it back into the right place.
To do so, we first adjust it's vruntime value by adding the time it ran and than use the insertion algorithm.
Due to the properties of red-black binary trees, all operations are O(log(n)) expensive.

\subsection{Priorities}

There isn't a special priority queue. The priority is adjusted by weighing the vruntime.
$$ \text{vruntime}_k = \text{vruntime}_{k-1} + t_{\text{ran}}*\text{weight} $$
If a process is more important, we adjust the weight, so that the time that it ran is counted as less. 
This means that even though it ran for the same amount of time as the other, it didn't receive the same "penalty".
Therefore it runs sooner than any other task.
In linux the weight is called "niceness" and the $t_{\text{ran}}$ is multiplied by $\frac{1}{\text{niceness}}$

\section*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}

In this paper we went from taking the first steps to trying to understand the real use application. 
I hoped to make it easy to get started by using relatable a analogy.
However, there is a point, where one has to abandon old beliefs to grasp newer ones.
As time went on the algorithms became more and more complex.

### 2024-12-31 LaTeX/Thesis.tex
added a bit more detail
added cfs and conclusion

          \section*{Further Sources}
          \begin{itemize}
            \item \url{https://github.com/Pseudomanifold/latex-mimosis}
            \item \url{https://en.wikipedia.org/wiki/Binary_search}
            \item \url{https://www.quora.com/Why-is-look-up-faster-for-a-Binary-Tree-than-a-Linked-List}
            \item \url{https://ceunican.github.io/aos/09.Scheduling_Proportional_Share.pdf}
            \item \url{https://courses.cs.washington.edu/courses/cse451/12au/l11.pdf}
          \end{itemize}

>>>>

          \section*{Further Sources}
          \begin{itemize}
            \item \url{https://github.com/Pseudomanifold/latex-mimosis}
            \item \url{https://en.wikipedia.org/wiki/Binary_search}
            \item \url{https://www.quora.com/Why-is-look-up-faster-for-a-Binary-Tree-than-a-Linked-List}
            \item \url{https://ceunican.github.io/aos/09.Scheduling_Proportional_Share.pdf}
            \item \url{https://courses.cs.washington.edu/courses/cse451/12au/l11.pdf}
            \item \url{https://en.wikipedia.org/wiki/Earliest_eligible_virtual_deadline_first_scheduling}
            \item \url{https://en.wikipedia.org/wiki/Proportional_share_scheduling}
            \item \url{https://www.youtube.com/watch?v=MkJfuI5_hjc&t=29s}
            \item \url{https://www.youtube.com/watch?v=qvZGUFHWChY&list=PL9xmBV_5YoZNqDI8qfOZgzbqahCUmUEin&index=1&pp=iAQB}
            \item \url{https://www.youtube.com/watch?v=95s3ndZRGbk&list=PL9xmBV_5YoZNqDI8qfOZgzbqahCUmUEin&index=2&pp=iAQB}
            \item \url{https://en.wikipedia.org/wiki/Red%E2%80%93black_tree} 
          \end{itemize}
